---
title: "Riddler ~ Jan 24, 2020"
author: "mleary"
date: "February, 21 2020"
output: 
  rmdformats::html_clean:
    highlight: kate
    collapsed: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```


URL: https://fivethirtyeight.com/features/can-you-flip-your-way-to-victory/

# Riddler Express: 

On a warm, sunny day, Nick glanced at a thermometer, and noticed something quite interesting. When he toggled between the Fahrenheit and Celsius scales, the digits of the temperature — when rounded to the nearest degree — had switched. For example, this works for a temperature of 61 degrees Fahrenheit, which corresponds to a temperature of 16 degrees Celsius.

However, the temperature that day was not 61 degrees Fahrenheit. What was the temperature?

## Approach:
### Setup

For this one I wanted to solve the problem in both R & Python, so I decided to use the reticulate package.

```{r}
#Setup
library(reticulate)
reticulate::use_python('/Library/Frameworks/Python.framework/Versions/3.8/bin/python3')

```

### R Solution

This was fun...

```{r}
library(tidyverse)
library(stringi)

# helper function to convert fahrenheit to celsius
# formula for conversion -- (2°C × 9/5) + 32 = 35.6°F
celsius_to_fahrenheit_round <- function(temp.celsius){
  temp.fahrenheit <-  (temp.celsius * (9/5)) + 32
  temp.fahrenheit <- round(temp.fahrenheit, digits = 0)
  return(temp.fahrenheit)
}

# create table
answer <-  tibble(temp.c = 1:100) %>% 
 mutate(temp.f = celsius_to_fahrenheit_round(temp.c),
        temp.c = as.character(temp.c),
        temp.f = as.character(temp.f),
        temp.f.switched = stringi::stri_reverse(temp.f)) %>% 
   filter(temp.c == temp.f.switched) %>% 
   filter(temp.c != 16) # filter out the record already mentioned
 
```

R answer is `r answer$temp.c` degres celsius or `r answer$temp.f` degrees fahrenheit.

### Python Solution

This wasn't the most elegant solution...

```{python}
#Setup
import pandas as pd
# helper function to convert fahrenheit to celsius
# formula for conversion -- (2°C × 9/5) + 32 = 35.6°F
def celsius_to_fahrenheit_round(tempcelsius):
    tempfahrenheit = (tempcelsius * (9/5)) + 32
    tempfahrenheit = round(tempfahrenheit )
    return tempfahrenheit

# helper function to reverse an integer
def revese_temp(temp):
    temp = str(temp)
    temp = list(temp)
    temp.reverse()
    temp = "".join(temp)
    temp = int(temp)
    return temp

tempC = pd.Series(range(1,101))
tempF = celsius_to_fahrenheit_round(tempC).astype(int)

tempDF = pd.DataFrame({"tempC":tempC, "tempF":tempF})

tempDF['tempFrev'] = tempDF.tempF.apply(revese_temp)

tempDF[tempDF.tempC == tempDF.tempFrev]

```

## Riddler Express Answer:
28 degrees celcius or 82 degress fahrenheit


